Robust Spectroscopic Inference with Imperfect Models

### 2014-09-10 tex/paper.tex
Appendix edits

\section{Interpolation Errors}
Interpolating a model spectrum from a synthetic library of spectra on a grid is conceptually straightforward; however, in practice there are several subtle effects to consider. Spectra are stored in a synthetic library as samples of the intensity at specific wavelengths, or pixels, at high resolution (typically R$\sim$500,000). Typically, to interpolate a spectrum for a specific.

Moreover, often the flux value of the pixel will not vary monontonically. Visually inspecting values of pixels as a function of parameter values makes it apparent that the spectra have not be Nyquist sampled in parameter space. That the synthetic grid does not contain the full amount of information about the spectral behavior. However, this is not surprising giving the computational complexity of stellar synthesis. Therefore, we are left to make the most conservative assumption about the behavior of the pixels between grid points and decide on a linear interpolation scheme. Going beyond this behavior, for example using a spline or other band-limited interpolation scheme, ignores the fact that the grid itself is not Nyquist sampled.

We choose to do linear interpolation, but empirically determine the interpolation error by a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, for example T=6000, logg=4.5, and Z=-0.5, and then create three sets of interpolated spectra, over the various dimensions. Therefore, for each spectrum in the synthetic grid, there are also three error spectra Generally, the interpolation error is small, typically no larger than 1\% across many lines. This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the multitude of possible lines to fit. However, while the interpolation error is small, it is often the case that the error adds \emph{coherently}. When the signal to noise of the spectrum is particularly large, this coherent trend can manifest itself as a behavior that the spectral lines in an interpolated spectrum are less ``physical'' looking than those directly near a gridpoint, with minor or no interpolation. Therefore, it is necessary to quantify the degree of interpolation error, in order to give the proper likelihood to interpolated spectra.

>>>>

\section{Interpolation Errors}
The interpolation of a model spectrum from a synthetic library of spectra is a conceptually straightforward process; however, in practice there may be several subtle effects that require consideration, depending on the application. Spectra are generally stored in a synthetic library as 1-D arrays of flux values, sampled at specific wavelengths at high resolution (typically R$\gtrsim$100,000). Each spectrum has several fundamental parameters which describe the model that generated it. To use a concrete example, the stellar libraries used in this paper have three parameters, $T_\textrm{eff}$, $\log_{10} g$, and $\Z$, each of which constitutes an axis in the synthetic library. Therefore, the full synthetic grid can be stored as a 4-D array, where the first three axes are the stellar parameters and the 4th axis describes the wavelength. The value stored in the array is the flux of the spectrum described by current value of the four axes.

Spectral line formation is a non-linear process, with the shape of the line determined by complicated radiative synthesis, which is a function of the stellar parameters. Interpolating a \emph{physical} spectrum short-cuts this non-linear process and attempts to predict the behavior of the spectral line based upon the behavior of nearby spectra in the grid. While this approximation is generally accurate enough for our applications, for high signal-to-noise applications it becomes necessary to account for any inaccuracies introduced by the interpolation. In practice, there are two factors which influence the choice of interpolation method. First, for a small percentage of pixels, the flux value will not vary monotonically with change in fundamental parameters. Second, visual inspection of the pixel values as a function of fundamental parameters makes it apparent that the spectra are not Nyquist sampled in parameter space. This means that synthetic grid does is not dense enough to losslessly reconstruct any synthetic spectrum. However, this fact is not surprising given the computational complexity of stellar synthesis. Given this reality, we make the most conservative assumption about the behavior of the pixels between grid points and decide to use a linear interpolation scheme. Using a more advanced interpolation method, such as a spline or other band-limited interpolation scheme, would not remedy the fact that the grid itself is not Nyquist sampled. To account for the loss of spectral information during the interpolation process, we build a model of interpolation error and incorporate this into the likelihood function (Eqn.~\ref{eqn:lnlikelihood}).

We empirically determine the interpolation error by performing a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, generate an interpolated spectrum from the nearest two grid points, and compare the difference between the two. For example, consider the {\sc Phoenix} grid used for the WASP-14 parameter estimation. In this parameter space, the grid is sampled every 100K in $T_\textrm{eff}$ and every 0.5 dex in $\log_{10} g$ and $\Z$. Therefore, for each spectrum in the synthetic grid, there are also three error spectra associated with it, one for each axis of stellar parameters. For example, consider a spectrum with stellar parameters $T_\textrm{eff} = $ 6000K, $\log_{10} g=$4.0 dex, and $\Z = $ -0.5 dex. A temperature ``error spectrum'' would be formed by subtracting this true spectrum from the spectrum formed by the average of the $T_\textrm{eff} = $ 6100K spectrum and the $T_\textrm{eff} = $ 5900K spectrum, with $\log_{10} g$ and $\Z$ held constant. Likewise, the $\log_{10} g$ and $\Z$ error spectra are created by performing the same operation over the $\log_{10} g$ and $\Z$ axes of the grid (see Figure~\ref{fig:interp_errors}).

These are done over a 200K range, and the actual range will only be 100K. Tests with other interpolation show that the interpolation error increases approximately linearly with interpolation distance. Therefore, when using the spectrum, we divide the total interpolation error spectrum by 2.

Generally, the interpolation error is small, typically no larger than 1\% across many lines. This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the multitude of possible lines to fit. However, while the interpolation error is small, it is often the case that the error adds \emph{coherently}. When the signal to noise of the spectrum is particularly large, this coherent trend can manifest itself as a behavior that the spectral lines in an interpolated spectrum are less ``physical'' looking than those directly near a gridpoint, with minor or no interpolation. Therefore, it is necessary to quantify the degree of interpolation error, in order to give the proper likelihood to interpolated spectra.
